 %Seção de acrônimos
\pretextualchapter{List of Acronyms}

\begin{description}
 \item BCET - Best-Case Execution Time .
 \item CP - Constraint Programming.
 \item GA - Genetic Algorithms.
 \item EMO - Evolutive Multi-objective algorithm.
 \item FOREPOST - Feedback-ORiEnted PerfOrmance Software Testing.
 \item Hybrid -Hybrid Metaheuristic approach proposed by this research.
 \item HybridQ - Hybrid Metaheuristic with Q-Learning approach proposed by this r.
 \item ICA - Independent Component Analysis.
 \item MBT - Model-based Testing.
 \item MDP - Markov Decision Process.
 \item MOEA/D - Multiobjective Evolutionary Algorithm Based on Decomposition.
 \item MOEAs - Multi-objective evolutionary algorithms. 
 \item MOPs - Multi-objective problems. 
 \item NIST -  National Institute of Standards and Technology. 
 \item NSGA II - Non-dominated Sorting Genetic Algorithm II.
 \item P-metaheuristics - Population-based metaheuristics.
 \item PAES - Pareto Archived Evolution Strategy.
 \item PASA - A Method for Performance Assessment of Software Architectures.
 \item PID - Proportional,Integral, and Derivative.
 \item S-metaheuristic - Single Solution or Single-objective metaheuristic.
 \item SA - Simulated Annealing.
 \item SBSE - Search-based Software Engineering.
 \item SBST - Search-based Software Testing. 
 \item SEDR - Standard Error Dynamic Resampling.
 \item SLAs - Service Level Agreements.
 \item SPEA2 - Strength Pareto Evolutionary Algorithm 2.
 \item TPC - Transaction Processing Performance.
 \item TS - Tabu Search.
 \item UCML -  User Community Modeling Language.
 \item WCET - Worst-Case Execution Time .

 
\end{description}
