%Seção de agradecimentos
\chapter*{Acknowledgments} 

%\begin{trivlist} \itemsep 2ex \normalsize


It has been an exciting and instructive study period at the University of Fortaleza and I feel privileged to have had the opportunity to carry out this study as a demonstration of
knowledge gained during the period. With these acknowledgments, it would be impossible to remember those who in one way or another, directly or indirectly, have played a role in the realization of this research project. 

First of all, Thank you God for the blessing and guiding me during all this project. I would like to dedicate this thesis to three of the most important people in my life
– my wife, my mother and my father – whose unconditional love and support gave me the
strength to follow my dream of becoming a scientist. I express all my gratitude to my beloved wife Najara, who endured five long years by my side without ever questioning my choice of becoming a doctoral degree student. Thank you to my parents, Francisco and Zirlanda,  for the example of life and for encouraging me as a child to enjoy science.

I would like to thank my very dedicated advisors Dr. Pedro Porfírio and Dr. André Coelho, for their guidance and support throughout this research project, which has been an excellent
learning experience for me. This thesis is not solely the product of my work but also the hard work and perseverance of my advisors, providing help to keep me going even in the most difficult of times to get the project done. I can only hope that more students will be given the opportunity to work and learn from both teachers as I did.

I also would like to thank the following people:

My thesis committee – Dr. Pedro de Alcantara dos Santos Neto, Dr. João Paulo Pordeus Gomes, Ph.D. Americo Tadeu Falcone Sampaio and Docteur Napoleão Vieira Nepomuceno – who made themselves available when I needed them to review previous drafts of this thesis and for their valuable suggestions that helped me to bring this work to its completion.

My family who loves me and is always present: my mother in law Aroldina, my brothers Nilton and Nelio, my nephew Rafael, my cousins, my uncles, my aunts and my grandmother Elisabete.

Dr. Adriano Bessa, Dr. Victor Hugo, Dr. José Aguiar Filho,  Dr. Plácido Pinheiro and all professors from the post-graduate program for their direct and indirect contributions.

My colleagues from UNIFOR who supported me during this project: Adriana, Marum Simão, Celso Medeiros and José Renato. 

My company, SERPRO — Brazilian Federal Datacenter, for the (partial) financial support. My colleagues Flavio Cysne, Fred Viana, Luis and Suderland Guimarães to help me with one of the experiments applied in this thesis. Sergio Gomes and Carlos Henrique (Managers of the Fortaleza Team) for giving me the time necessary to develop my activities.

Julio Tavares, Carlos Manso and Marcia Tonieto. Coordinators of the Computer Science Course at
Faculdade Metropolitana da Grande Fortaleza - Fametro and Faculdade Lourenço Filho - FLF. For all your support over the years that I taught in those colleges. My students from the various disciplines that I taught.

